---
title: "Instrumental Variables"
date: "null"
editor_options:
  chunk_output_type: console
output:
  github_document:
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
always_allow_html: yes
---

```{r setup, echo = F, message = F}
knitr::opts_chunk$set(
  cache = T,
  cache.path = '../cache/',
  fig.path = '../fig/',
  message = F,
  warning = F
  )

library(hrbrthemes)
library(mixtape)
library(stargazer)
library(estimatr)
library(broom)
library(dplyr)
library(car)
library(sandwich)
library(lmtest)

```

This exercise uses

- `dplyr` for data wrangling 
- `broom::tidy()` for extracting coefficient estimates and standard errors
- `estimatr::lm_robust()` for linear regression with robust standard errors 
- `estimatr::iv_robust()` for 2SLS with robust standard errors 
- `car::linearHypothesis()` to run conditional F test for weak instruments 


# College in the county

## First-stage 

Regress the endogenous covariate on all exogenous variables

```{r}
model2 <- lm_robust(educ ~ nearc4 + exper + black + south + married + smsa,
                    data = card)

tidy(model2) %>% 
  mutate_at(vars(estimate, std.error), .funs = ~ round(., 3))

linearHypothesis(model2, "nearc4 = 0")
```

## Estimate OLS and 2SLS effects 

The formula syntax for `estimatr::iv_robust()``` is: 

```r
iv_robust(Y ~ D + X | Z + X, data = dat)
```

where $D$ is the endogenous variable of interest, $X$ are controls and $Z$ is the instrument. 

```{r}
model1 <- lm_robust(lwage ~ educ + exper + black + south + married + smsa, 
                    data = card) 
  
model1 %>% 
  tidy() %>% 
  mutate_at(vars(estimate, std.error), .funs = ~ round(., 3))

ivreg <- iv_robust(lwage ~ educ + exper + black + south + married +
            smsa  | nearc4 + exper + black + south + married + smsa,
          data = card, se_type = "classical") 
```

# Fulton fish markets

## OLS

```{r}

lm_robust(log(quantity) ~ log(price) + mon + tues + wed + thurs + time, 
          data = fish) %>% 
  tidy()
```

## 2SLS with average wave height instrument 

```{r}
fs_waveheight <- lm_robust(log(price) ~ wave2 + mon + tues + wed + thurs + time, 
                           data = fish)

fs_waveheight %>% 
  tidy() %>% 
  mutate_at(vars(estimate, std.error), .funs = ~ round(., 3))

linearHypothesis(fs_waveheight, "wave2 = 0")
```

```{r}
iv_robust(log(quantity) ~ log(price) + mon + tues + wed + thurs + time | 
            wave2 + mon + tues + wed + thurs + time, 
          data = fish) %>% 
  tidy() %>% 
  mutate_at(vars(estimate, std.error), .funs = ~ round(., 3))
```

## 2SLS with wind speed instrument 

### First-stage
```{r}
fs_waveheight <- lm_robust(log(price) ~ speed3 + mon + tues + wed + thurs + time, 
                           data = fish)

fs_waveheight %>% 
  tidy() %>% 
  mutate_at(vars(estimate, std.error), .funs = ~ round(., 3))

linearHypothesis(fs_waveheight, "speed3 = 0")
```

### 2SLS

```{r}
iv_robust(log(quantity) ~ log(price) + mon + tues + wed + thurs + time | 
            speed3 + mon + tues + wed + thurs + time , 
          data = fish) %>% 
  tidy() %>% 
  mutate_at(vars(estimate, std.error), .funs = ~ round(., 3))
```

