---
title: "Instrumental Variables"
date: "null"
editor_options:
  chunk_output_type: console
output:
  github_document:
    
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
always_allow_html: yes
---

```{r setup, echo = F, message = F}
knitr::opts_chunk$set(
	fig.path = "../fig/",
	message = FALSE,
	warning = FALSE
)

library(hrbrthemes)
library(mixtape)
library(stargazer)
library(estimatr)
library(broom)
library(dplyr)
library(car)
library(sandwich)
library(lmtest)

# short function to round and keep only important parts of the output
tidy_round <- . %>% 
  tidy() %>% 
  select(term, estimate, std.error, p.value) %>%
  knitr::kable(digits = 3) %>% 
  kableExtra::kable_styling()

```

This exercise uses

- `dplyr` for data wrangling 
- `broom::tidy()` for extracting coefficient estimates and standard errors
- `estimatr::lm_robust()` for linear regression with robust standard errors 
- `estimatr::iv_robust()` for 2SLS with robust standard errors 
- `car::linearHypothesis()` to run conditional F test for weak instruments 


# College in the county

## First-stage 

Regress the endogenous covariate on all exogenous variables

```{r}
model2 <- lm_robust(educ ~ nearc4 + exper + 
                      black + south + married + smsa,
                    data = card)

tidy_round(model2)

linearHypothesis(model2, "nearc4 = 0")
```

## Estimate OLS and 2SLS effects 

The formula syntax for `estimatr::iv_robust()` is

```r
iv_robust(Y ~ D + X | Z + X, data = dat)
```

where $D$ is the endogenous variable of interest, $X$ are controls and $Z$ is the instrument. 

<table>
<tr>
<td>
```{r}
lm_robust(lwage ~ educ + exper +
            black + south + married + smsa,
          data = card) -> model1
  
tidy_round(model1)
```
</td>
<td>
```{r}
iv_robust(lwage ~ educ + exper +
            black + south + married +
            smsa  | nearc4 + exper +
            black + south + married + smsa,
          data = card, 
          se_type = "classical") %>%  
  tidy_round()
```
</td>
</tr>
</table>

# Fulton fish markets

</table>
<tr>
<td>


```{r}

lm_robust(log(quantity) ~ log(price) + 
            mon + tues + wed + thurs + time, 
          data = fish) %>% 
  tidy_round()
```

</td>
<td>


```{r}
fs_waveheight <- lm_robust(log(price) ~ wave2 + 
                             mon + tues + wed + thurs + time, 
                           data = fish)

tidy_round(fs_waveheight)
```

</td>
</tr>
</table>

```{r}
linearHypothesis(fs_waveheight, "wave2 = 0")
```

```{r}
iv_robust(log(quantity) ~ log(price) + mon + tues + wed + thurs + time | 
            wave2 + mon + tues + wed + thurs + time, 
          data = fish, se_type = "classical") %>% 
  tidy_round()
```

## 2SLS with wind speed instrument 

<table>
<tr>
<td>

### First-stage
```{r}
lm_robust(log(price) ~ speed3 + mon +
            tues + wed + thurs + time,
          data = fish) -> fs_waveheight

tidy_round(fs_waveheight)

```

</td>
<td>

### 2SLS

```{r}
iv_robust(log(quantity) ~ log(price) + mon + tues + 
            wed + thurs + time | 
            speed3 + mon + tues + wed + thurs + time , 
          data = fish, se_type = "classical") %>% 
  tidy_round()
```

</td>
</tr>
</table>

```{r}
linearHypothesis(fs_waveheight, "speed3 = 0")
```
